Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-09-04 papers/thesis-kirstie/thesis.tex
updated more figure references
updated another figure reference
modified discussion on B2 and moved stuff to appendix

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{plot_HardSphere_Potential.png}
    \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity for $r\leq{2R}$ which is approximately 1.225 in this example.}
    \label{fig:HardSphere_potential}
  \end{figure}
\[{}\]
\[{}\]
\[{}\]
\[{}\]

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{LJ_Potential.png}
    \caption{Leonard-Jones Potential for $\sigma=1$, $\epsilon=1$}
    \label{fig:LJ_potential}
  \end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{LJ_Potential_2part.png}
    \caption{Leanard-Jones Potential broken into two parts for $\sigma=1$, $\epsilon=1$}
    \label{fig:LJ_potential_2parts}
  \end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{WCA_Potential.png}
    \caption{WCA Potential for $\sigma=1$, $\epsilon=1$}
    \label{fig:WCA_potential}
  \end{figure}

@@

\begin{figure}[h!]
    \centering
   % \begin{subfigure}[b]{0.4\linewidth}
      % FIXME \includegraphics[height=3.5cm]{TwoSpheres.png}
    %\end{subfigure} 
    %\begin{subfigure}[b]{0.4\linewidth}
    %  \includegraphics[height=3.5cm]{2spheres.png}
    %\end{subfigure} 
    \caption{(Left) The WCA potential is a minimum at r=2R. 
             (Right) The WCA potential increases rapidly for $r<2R$.}
    \label{fig:TwoSpheres}
    \end{figure} 
%\[\]

The WCA potential is not well suited for implementing computationally.  Therefore, an erf potential which closely approximates the WCA potential at temperatures of interest will be used instead of the WCA potential. The erf potential is given by: \begin{equation}{V_{erf}=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}\end{equation} where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA potentail when r equals $\alpha$ \begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} and is given by \begin{equation}{\alpha=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}\end{equation}  The parameter $\Xi$ is derived by matching the slope of the erf potential to the slope of the WCA potential at r=$\alpha$:\begin{displaymath}{\frac{dV_{erf}(r=\alpha)}{dr}=\frac{dV_{WCA}(r=\alpha)}{dr}}\end{displaymath} and is given by: \begin{equation}{\Xi=\frac{\alpha}{6\sqrt{\pi}\left(\sqrt{\frac{\epsilon}{k_BT}\ln(2)}+\ln(2)\right)}}\end{equation}

@@

 \begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.4\linewidth}
      % FIXME \includegraphics[height=4.5cm]{plot_alpha.png}
      \includegraphics[height=5cm]{plot_alpha.pdf}
      \caption{$\alpha$ versus $k_BT/\epsilon$}
    \end{subfigure} 
    \begin{subfigure}[b]{0.4\linewidth}
      % FIXME \includegraphics[height=4.5cm]{plot_xi.png}
      \includegraphics[height=5cm]{plot_xi.pdf}
      \caption{$\Xi$ versus $k_BT/\epsilon$}
    \end{subfigure} 
    \caption{These plots, generated with $\sigma=1$ and $\epsilon=1$, show how the parameters $\alpha$ and $\Xi$ vary with temperature. (a) $\alpha$ decreases with increasing temperature. (b) $\Xi$ rapidly increases for temperatures below $\frac{k_BT}{\epsilon}=52$ and then gradually decreases.}
    \label{fig:alphaXivsT}
    \end{figure}

@@

Another way to find values for $\Xi(T)$ is to find computationally the value $\Xi$ for which the second virial coefficient evaluated for the error function potential $V_{erf}(r)$ is equal to the value for the second virial coefficent evalued for the WCA potential $V_{WCA}$, with $\alpha(T)$ plugged into the equation for $B_{2erf}$ at each temperature T. 
\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
\begin{equation}B_{2erf}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{erf}(r)}-1\right)d\vec{r} \end{equation}
\begin{equation}B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]\end{equation}
and 
\begin{equation}B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} \end{equation}
A plot of $\Xi(T)$ obtained through this method is shown in Figure \ref{fig:xi_fromB2vsT}, and Figure \ref{fig:erf_Potential_xifromB2} shows the erf potential at reduced temperatures $\frac{k_BT}{\epsilon}=1$, $\frac{k_BT}{\epsilon}=2$, and $\frac{k_BT}{\epsilon}=3$ alongside the WCA potential.

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{xi_fromB2.pdf}
    \caption{Shows the temperature dependence of $\Xi$ generated from matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ at each temperature.}
    \label{fig:xi_fromB2vsT}
  \end{figure}
\color{red}

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{erf_Potential_xifromB2.pdf}
    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
    \label{fig:erf_Potential_xifromB2}
  \end{figure}

\noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
\color{black}
%\[\]
%\[\]

%----------------BEGIN Virial Expansion--------------
The second virial coefficient can be derived from Euler's equation for the internal energy of a system.
\begin{equation}PV=TS-(U-\mu{N})\end{equation}
\begin{equation}\label{Euler_rearranged}\frac{PV}{k_BT}=\frac{S}{k_B}-\frac{1}{k_BT}(U-\mu{N})\end{equation}
Setting $\beta=\frac{1}{k_BT}$, and putting in the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
%\begin{equation}\beta=\frac{1}{k_BT}\end{equation}
%\begin{equation}\frac{PV}{k_BT}=-\sum{P_i\ln{P_i}}-\frac{1}{k_BT}(\sum{P_iE_i}-\mu\sum{P_iN})\end{equation}
Equation~\ref{Euler_rearranged} then becomes
\begin{equation}\frac{PV}{k_BT}=-\sum{P_i\ln{P_i}}-\sum{P_i\beta(E_i-\mu
{n_i}})\end{equation}
\begin{equation}\frac{PV}{k_BT}=-\sum{P_i}\left[\ln{P_i}+\beta(E_i-\mu
{n_i})\right]\end{equation}
For a Grand Canonical Ensemble the probablity of a state given by
\begin{equation}P_{n,s(n)}=\frac{e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\end{equation}
results in 
\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[\ln{\left(\frac{ e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}

%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength is smaller than the mean distance between the spheres. Thus, the WCA fluid can be treated as a classical fluid.
%\\

%\[\]
%\[\]
\[\]
\noindent IB. The Theory of Freezing and Maxwell's Double Tangent Construction

@@

\begin{figure}[h!]
    \centering
    % FIXME \includegraphics[height=7cm]{MDTC.png}
    \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

@@

The pressure at the transistion point can be obtained from the point of intersection that occurs when the Gibbs free energy per atom is plotted against the pressure for both the liquid and the solid at fixed temperature, as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from the slope and cooresponding y-intercept data collected from lines tangent along the curves on the the Helmholtz free energy per atom verses inverse density plots for both the liquid and the solid. 
\begin{figure}[h!]
    \centering
    % FIXME \includegraphics[height=7cm]{MDTC-2.png}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

@@

\begin{figure}[h!]
    \centering
    % FIXME \includegraphics[height=7cm]{MDTC-3.png}
    \caption{The liquid and solid densities at the point of transistion are found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

@@

 \begin{figure}[h!]
    \centering
    %\includegraphics[height=4cm]{HomogeneousDen.png}
    % FIXME \includegraphics[height=4cm]{Homogeneous_bold-box.png}
    \caption{A homogeneouse number density profile $n(\vec{r})$}
    \label{fig:homogen_denisty}
  \end{figure}

@@

>>>>

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{plot_HardSphere_Potential.pdf}
    \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity for $r\leq{2R}$ which is approximately 1.225 in this example.}
    \label{fig:HardSphere_potential}
  \end{figure}
\[{}\]
\[{}\]
\[{}\]
\[{}\]

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{LJ_Potential.pdf}
    \caption{Leonard-Jones Potential for $\sigma=1$, $\epsilon=1$}
    \label{fig:LJ_potential}
  \end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{LJ_Potential_2part.pdf}
    \caption{Leanard-Jones Potential broken into two parts for $\sigma=1$, $\epsilon=1$}
    \label{fig:LJ_potential_2parts}
  \end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{WCA_Potential.pdf}
    \caption{WCA Potential for $\sigma=1$, $\epsilon=1$}
    \label{fig:WCA_potential}
  \end{figure}

@@

\begin{figure}[h!]
    \centering
      % FIXME \includegraphics[height=3.5cm]{TwoSpheres.pdf} 
    \caption{(Left) The WCA potential is a minimum at r=2R. 
             (Right) The WCA potential increases rapidly for $r<2R$.}
    \label{fig:TwoSpheres}
    \end{figure}

The WCA potential is not well suited for implementing Density Functional Theory.  Therefore, an erf potential which closely approximates the WCA potential at temperatures of interest will be used instead of the WCA potential. The erf potential is given by 
\begin{equation}
	{V_{erf}=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}
\end{equation} 
where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA potentail when r equals $\alpha$ 
\begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} 
and is given by 
\begin{equation}\label{alphaT}
	{\alpha=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
\end{equation} 
The parameter $\Xi$ is derived by matching the slope of the erf potential to the slope of the WCA potential at r=$\alpha$:
\begin{displaymath}{\frac{dV_{erf}(r=\alpha)}{dr}=\frac{dV_{WCA}(r=\alpha)}{dr}}\end{displaymath} 
and is given by: 
\begin{equation}
	{\Xi=\frac{\alpha}{6\sqrt{\pi}\left(\sqrt{\frac{\epsilon}{k_BT}\ln(2)}+\ln(2)\right)}}
\end{equation}

@@

 \begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.4\linewidth}
      \includegraphics[height=5cm]{plot_alpha.pdf}
      \caption{$\alpha$ versus $k_BT/\epsilon$}
    \end{subfigure} 
    \begin{subfigure}[b]{0.4\linewidth}
      \includegraphics[height=5cm]{plot_xi.pdf}
      \caption{$\Xi$ versus $k_BT/\epsilon$}
    \end{subfigure} 
    \caption{These plots, generated with $\sigma=1$ and $\epsilon=1$, show how the parameters $\alpha$ and $\Xi$ vary with temperature. (a) $\alpha$ decreases with increasing temperature. (b) $\Xi$ rapidly increases for temperatures below $\frac{k_BT}{\epsilon}=52$ and then gradually decreases.}
    \label{fig:alphaXivsT}
    \end{figure}

@@

Another way to find values for $\Xi(T)$ is to use the second virial coefficent $B_{2}$ from the equation of state expressed in terms of the virial expansion. 
\begin{equation}\label{virialequation}
    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
\end{equation}
where 
\begin{equation}
	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
\end{equation}
The second virial coefficient can be derived from Euler's equation for the internal energy of a system, 
\begin{equation}U=TS-PV+\mu{N}\end{equation}
using the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
which gives
\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
Expressing the Helmholtz free energy in terms of the pair potential $V(r)$ results in the equation of state expressed in terms of the virial expansion as given by Equation~\ref{virialequation}.

The value of $\Xi$ at a given temperature, T, can be found computationally by equating the second virial coefficient evaluated for the error function potential $V_{erf}(r)$, and the second virial coefficent evalued for the WCA potential $V_{WCA}$.
\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
Evaluating $B_{2}$ for the erf potential gives
%\begin{equation}B_{2erf}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{erf}(r)}-1\right)d\vec{r} \end{equation}
\begin{equation}
	B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
\end{equation}
where $\alpha(T)$ is given by Equation~\ref{alphaT} as before. Similarly, $B_{2WCA}$ can be found computationally from 
\begin{equation}
	B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} 
\end{equation}

A plot of $\Xi(T)$ obtained through this method is shown in Figure \ref{fig:xi_fromB2vsT}, and Figure \ref{fig:erf_Potential_xifromB2} shows the erf potential at a few different temperatures alongside the WCA potential.

\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{xi_fromB2.pdf}
    \caption{Shows the temperature dependence of $\Xi$ generated from matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ at each temperature.}
    \label{fig:xi_fromB2vsT}
  \end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{erf_Potential_xifromB2.pdf}
    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
    \label{fig:erf_Potential_xifromB2}
  \end{figure}

%\color{red}
% FIXME \noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
%\color{black}

%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength is smaller than the mean distance between the spheres. Thus, the WCA fluid can be treated as a classical fluid.

\[\]
\noindent IB. The Theory of Freezing and Maxwell's Double Tangent Construction

@@

\begin{figure}[h!]
    \centering
     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig1.pdf}
    \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

@@

The pressure at the transistion point can be obtained from the point of intersection that occurs when the Gibbs free energy per atom is plotted against the pressure for both the liquid and the solid at fixed temperature, as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from the slope and cooresponding y-intercept data collected from lines tangent along the curves on the the Helmholtz free energy per atom verses inverse density plots for both the liquid and the solid. 
\begin{figure}[h!]
    \centering
     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig2.pdf}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

@@

\begin{figure}[h!]
    \centering
     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig3.pdf}
    \caption{The liquid and solid densities at the point of transistion are found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

@@

 \begin{figure}[h!]
    \centering
    %\includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
    % FIXME \includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
    \caption{A homogeneouse number density profile $n(\vec{r})$}
    \label{fig:homogen_denisty}
  \end{figure}

@@

\noindent \textbf{Derivation of the Virial Equation}
%----------------BEGIN Virial Expansion--------------
The second virial coefficient can be derived from Euler's equation for the internal energy of a system.
\begin{equation}PV=TS-(U-\mu{N})\end{equation}
\begin{equation}\label{Euler_rearranged}\frac{PV}{k_BT}=\frac{S}{k_B}-\frac{1}{k_BT}(U-\mu{N})\end{equation}
Setting $\beta=\frac{1}{k_BT}$, and putting in the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
%\begin{equation}\beta=\frac{1}{k_BT}\end{equation}
%\begin{equation}\frac{PV}{k_BT}=-\sum{P_i\ln{P_i}}-\frac{1}{k_BT}(\sum{P_iE_i}-\mu\sum{P_iN})\end{equation}
Equation~\ref{Euler_rearranged} then becomes
\begin{equation}\frac{PV}{k_BT}=-\sum{P_i\ln{P_i}}-\sum{P_i\beta(E_i-\mu
{n_i}})\end{equation}
\begin{equation}\frac{PV}{k_BT}=-\sum{P_i}\left[\ln{P_i}+\beta(E_i-\mu
{n_i})\right]\end{equation}
For a Grand Canonical Ensemble the probablity of a state given by
\begin{equation}P_{n,s(n)}=\frac{e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\end{equation}
results in 
\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[\ln{\left(\frac{ e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}
